### 2020-06-18 TCS_version/06b-bisimulation.tex
reviewer 1

% !TEX root = ./main_RS2L_LNCS.tex
\section{Bio-simulation}
\label{sec:biosimulation}

The classical notion of bisimulation for process algebras equates two processes
when one process can simulate all the instructions executed by the other one and viceversa.
In its weak formulation, internal instructions, i.e. non visible by external observers,
are abstracted away. 
There are many variants of the bisimulation for process algebras,  for example the
barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when the expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
behavioural equivalence equates two processes when they expose the same ambients~\cite{GC03}.

@@

For instance we would like to express properties about each step of the simulation of a system like in the following:

\begin{example} \label{properties_ex}
\begin{enumerate}
\item Has the entity $s_i$ been used by rule $r_j$ as reagent?
\item Has the entity $s_i$ blocked the application of  rule $r_j$?
\item Has the entity $s_i$ been produced by rule $r_j$?
\item Has the entity $s_i$ been produced by some rule?
\item Has the entity $s_i$ been provided by the context? 
\item Has the rule $r_j$ not been applied? 
%\item Has the entity $s$ been used by some rule as reagent ?
%\item Has the entity $s$ blocked  the application of some rules ?
%\item Has the entity $s$ been produced ?
%\item Has the entity $s$ been provided by the context ?
%\item Has the rule $i$ been applied ?
\end{enumerate}
\end{example}

@@

Next, we introduce an assertion language that operates on simple strings
and that combines regular expression operators with conjunction and disjunction.
In the following we let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.

@@

The formulas corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:

\begin{enumerate}
\item
$\star :: r_j :: [ s_1,...,s_n ]^* :: s_i :: [  s_1,...,s_n ]^* :: [  \overline{s}_1,...,\overline{s}_n ]^+ :: \star$
\item
$\star :: r_j :: [s_i, \overline{s}_i] :: r_{j+1} :: \star$
\item
$\star :: p_j :: [  \mathit{ents} ]^* :: \widetilde{s}_i :: \star$
\item
$\star :: \widetilde{s}_i :: \star$
\item
$\star :: \widehat{s}_i :: \star$
\item
$\star :: r_j :: ? :: r_{j+1} :: \star$
\end{enumerate}
where in 1, 2, 6 we exploit the fact that in a reaction $(R,I,P)$ the sets $R$ of reactants and  $I$ of inhibitors are non empty, so that if there is only one symbol between the occurrence of $r_j$ and $r_{j+1}$ it means the reaction $r_j$ has not been applied. Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.

@@

The notion of bio-simulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
Please recall that we have defined the behaviour of the contest in a non deterministic way, thus 
at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
Bio-simulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.

@@

\begin{definition}[Bio-similarity  $ \sim_{\mathsf{F}}$]
Given an assertion $\mathsf{F}$, a \emph{bio-simulation} $\mathbf{R}_{\mathsf{F}}$ that respects $\mathsf{F}$ is a binary relation over c\CNA \  processes such that, whenever $P~\mathrel{\mathbf{R}_{\mathsf{F}}}~Q$ then:
 \begin{itemize} 
 \item
 for any $v,P'$ such that 
 $P \xRightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xRightarrow{w} Q'$ with $v\equiv_{\mathsf{F}} w$ and $P' \mathrel{\mathbf{R}_{\mathsf{F}}} Q'$.
\item
 for any $w,Q'$ such that 
 $Q \xRightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xRightarrow{v} P'$ with $v\equiv_{\mathsf{F}} w$ and $P' \mathrel{\mathbf{R}_{\mathsf{F}}} Q'$.
\end{itemize}
%\item i$Q \xrightarrow{v} Q'$ with $v \entails F$ iff $\exists v',P'$ s.t.~$v \stretcheq v'$, $P  \xrightarrow{v'} P'$ with $v \entails F$. and $P' \mathrel{\mathbf{R}_{F}} Q'$.
%\end{itemize}
We let $ \sim_{\mathsf{F}} $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $\mathsf{F}$, if $P \sim_{\mathsf{F}} Q$.
\end{definition}

\begin{remark}
Please remember that the notation $P \xRightarrow{v} P'$ refers to ordinary transitions $P \xrightarrow{v} P'$ where $v$ is a complete chain (solid and with $\silent$ actions at the extremes).
It should not be confused with the notation for weak transitions commonly found in the literature on process algebras.
\end{remark}

@@

\begin{definition}[Semantics of BioHML]
 We define $\llbracket \mathsf{G}\rrbracket \subseteq {\cal P}$ by induction on the structure of $\mathsf{G}$:
 %(read as the transition label $v$ satisfies the formula $F$) as: 
\[
\begin{array}{ccl@{\hspace{1cm}}ccl}
 \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket \mathsf{G}\wedge \mathsf{H}\rrbracket &\defeq&  \llbracket \mathsf{G}\rrbracket  \cap \llbracket \mathsf{H}\rrbracket  \\
\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
  \llbracket \mathsf{G}\vee \mathsf{H}\rrbracket &\defeq&  \llbracket \mathsf{G}\rrbracket  \cup \llbracket \mathsf{H}\rrbracket \\
%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
\end{array}
\]
\[
\begin{array}{lcl}
 \llbracket  \langle \chi \rangle \mathsf{G} \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xRightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  \mathsf{G} \rrbracket\} \\
 \llbracket [\chi]\mathsf{G} \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xRightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  \mathsf{G} \rrbracket\}
\end{array}
\]
We write $P \entails \mathsf{G}$ ($P$ satisfies $\mathsf{G}$) if and only if $P \in  \llbracket \mathsf{G} \rrbracket$.
\end{definition}

>>>>

% !TEX root = ./main_TCS.tex
\section{Bio-simulation}
\label{sec:biosimulation}

The classical notion of bisimulation for process algebras equates two processes
when one process can simulate all the instructions executed by the other one and viceversa.
In its weak formulation, internal instructions, i.e. non visible by external observers,
are abstracted away. 
There are many variants of the bisimulation for process algebras,  for example the
barbed bisimulation~\cite{10.1007/3-540-55719-9_114} only consider the execution of invisible actions, and then equates two processes when they expose the same prefixes; for the mobile ambients~\cite{CardelliG00}, a process algebra equipped with a reduction semantics, a notion of
behavioural equivalence equates two processes when they expose the same ambients~\cite{GC03}.

@@

\begin{example} \label{properties_ex}
For instance we would like to express properties about each step of the bio-simulation of a system like the ones below:
\begin{enumerate}
\item Has the entity $s_i$ been used by rule $r_j$ as reagent?
\item Has the entity $s_i$ blocked the application of  rule $r_j$?
\item Has the entity $s_i$ been produced by rule $r_j$?
\item Has the entity $s_i$ been produced by some rule?
\item Has the entity $s_i$ been provided by the context? 
\item Has the rule $r_j$ not been applied? 
%\item Has the entity $s$ been used by some rule as reagent ?
%\item Has the entity $s$ blocked  the application of some rules ?
%\item Has the entity $s$ been produced ?
%\item Has the entity $s$ been provided by the context ?
%\item Has the rule $i$ been applied ?
\end{enumerate}
\end{example}

@@

Next, we introduce an assertion language that operates on simple strings
and that combines regular expression operators with conjunction and disjunction.
We let $\mathit{syms} = \mathit{decs}\cup\mathit{reacts}\cup\mathit{prods}\cup\{\mathit{cxt}\}$ be the set of symbolic names used in our assertion language.

@@

\begin{example}
The formulas corresponding to the sample queries listed in Example~\ref{properties_ex} are as follows:
\begin{enumerate}
\item
$\star :: r_j :: [ s_1,...,s_n ]^* :: s_i :: [  s_1,...,s_n ]^* :: [  \overline{s}_1,...,\overline{s}_n ]^+ :: \star$
\item
$\star :: r_j :: [s_i, \overline{s}_i] :: r_{j+1} :: \star$
\item
$\star :: p_j :: [  \mathit{ents} ]^* :: \widetilde{s}_i :: \star$
\item
$\star :: \widetilde{s}_i :: \star$
\item
$\star :: \widehat{s}_i :: \star$
\item
$\star :: r_j :: ? :: r_{j+1} :: \star$
\end{enumerate}
where in 1, 2, 6 we exploit the fact that in a reaction $(R,I,P)$ the sets $R$ of reactants and  $I$ of inhibitors are non empty, so that if there is only one symbol between the occurrence of $r_j$ and $r_{j+1}$ it means the reaction $r_j$ has not been applied. Viceversa, if the reaction $r_j$ has been applied the occurrence of $r_j$ must be followed by at least one of the symbols in $\{s_1,...,s_n\}$ and then by at least one of the symbols in $\{\overline{s}_1,...,\overline{s}_n\}$.
\end{example}

@@

The notion of bio-simulation builds on the above language of assertions to parameterize the induced equivalence on the property of interest. 
Please recall that we have defined the behaviour of the context in a non deterministic way, thus 
at each step, different possible sets of entities can be provided to the system and different sets of reaction can be enabled/disabled. 
Bio-simulation can thus be used to compare the behaviour of different systems that share some of the reactions or entities or also to compare the behaviour of the same set of reaction rules when different contexts are provided.

@@

\begin{definition}[Bio-similarity  $ \sim_{\mathsf{F}}$]
Given an assertion $\mathsf{F}$, a \emph{bio-simulation} $\mathbf{R}_{\mathsf{F}}$ that respects $\mathsf{F}$ is a binary relation over c\CNA \  processes such that, whenever $P~\mathrel{\mathbf{R}_{\mathsf{F}}}~Q$ then:
 \begin{itemize} 
 \item
 for any $v,P'$ such that 
 $P \xRightarrow{v} P'$ then there exist $w,Q'$ such that $Q  \xRightarrow{w} Q'$ with $v\equiv_{\mathsf{F}} w$ and $P' \mathrel{\mathbf{R}_{\mathsf{F}}} Q'$.
\item
 for any $w,Q'$ such that 
 $Q \xRightarrow{w} Q'$ then there exist $v,P'$ such that $P  \xRightarrow{v} P'$ with $v\equiv_{\mathsf{F}} w$ and $P' \mathrel{\mathbf{R}_{\mathsf{F}}} Q'$.
\end{itemize}
%\item i$Q \xrightarrow{v} Q'$ with $v \entails F$ iff $\exists v',P'$ s.t.~$v \stretcheq v'$, $P  \xrightarrow{v'} P'$ with $v \entails F$. and $P' \mathrel{\mathbf{R}_{F}} Q'$.
%\end{itemize}

We let $ \sim_{\mathsf{F}} $ denote the largest  bio-simulation and we say that $P$ is \emph{bio-similar} to $Q$, with respect to $\mathsf{F}$, if $P \sim_{\mathsf{F}} Q$.
\end{definition}

\begin{remark}
Please remember that the notation $P \xRightarrow{v} P'$ refers to ordinary transitions $P \xrightarrow{v} P'$ where $v$ is a complete chain (solid and with $\silent$ actions at the extremes).
The double arrow notation should not be confused with the notation for weak transitions commonly found in the literature on process algebras.
\end{remark}

@@

\begin{definition}[Semantics of BioHML]
 We define $\llbracket \mathsf{G}\rrbracket \subseteq {\cal P}$ by induction on the structure of $\mathsf{G}$:
 %(read as the transition label $v$ satisfies the formula $F$) as: 
\[
\begin{array}{ccl@{\hspace{1cm}}ccl}
 \llbracket {\tt t}\rrbracket &\defeq& {\cal P} &  \llbracket \mathsf{G}\wedge \mathsf{H}\rrbracket &\defeq&  \llbracket \mathsf{G}\rrbracket  \cap \llbracket \mathsf{H}\rrbracket  \\
\llbracket {\tt f}\rrbracket& \defeq& \emptyset &
  \llbracket \mathsf{G}\vee \mathsf{H}\rrbracket &\defeq&  \llbracket \mathsf{G}\rrbracket  \cup \llbracket \mathsf{H}\rrbracket \\
%\item   $ v \entails F_1 \vee F_2$ iff   $v \entails F_1$  $\vee$  $v \entails F_2$.
\end{array}
\]
\[
\begin{array}{lcl}
 \llbracket  \langle \chi \rangle \mathsf{G} \rrbracket &\defeq& \{P \in {\cal P}: \exists v,P'.\ P\xRightarrow{v}P' \mbox{ with } v \entails \chi \mbox{ and } P' \in  \llbracket  \mathsf{G} \rrbracket\} \\
 \llbracket [\chi]\mathsf{G} \rrbracket &\defeq& \{P \in {\cal P}: \forall v,P'.\ P\xRightarrow{v}P' \mbox{ implies } v \entails \chi \mbox{ and } P' \in  \llbracket  \mathsf{G} \rrbracket\}
\end{array}
\]

We write $P \entails \mathsf{G}$ ($P$ satisfies $\mathsf{G}$) if and only if $P \in  \llbracket \mathsf{G} \rrbracket$.
\end{definition}

### 2020-06-18 TCS_version/09-appendix.tex
reviewer 1

 %\begin{proposition}(correctness 1)\\
 %\label{lem:struct}
%  Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be a $n$-step interactive process in $A$, with
% $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$.
%% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
%% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
%  Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation, and let $P'$ be such that $P \xrightarrow{\link{\silent}{\silent}}P'$.% then $P' = [|{\cal A},\gamma' |]$. 
%Then it holds that\\
% $P= \restrict{reacts,ctxs,ents} (\Pi_{a \in A}P_a| \Pi_{s \in S}CXT_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$\\
% $P'= \restrict{reacts,ctxs,ents} (\Pi_{a \in A}P_a| \Pi_{s \in S}CXT_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$
 %\end{proposition}
 \begin{proof}
 First, we note that all the channels in the system are restricted, see Def.~\ref{def:trans}, then
 it holds that the transition labels are of the form  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$.\\
 Now,
by Definition~\ref{def:trans} and by Lemma~\ref{lem:struct}.1, all the  channels $r_j,p_j$, with
 $j \in [1,\dots,u]$, and $cxt_h$, with $h \in [1, \dots, w]$, and all the annotated versions of 
 $s_i$, $s_o$ are restricted. Also, processes $\mathit{Cxt}$ always requires the 
 interaction with $P_s$ on either on channels $\widehat{s}_i$, $\widehat{s}_o$ or on channels 
 $\underline{s}_i$, $\underline{s}_o$.
 It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $\mathit{Cxt}$ (coding the behaviour of the context regarding all the entities) have been involved in the transition.

For any process $P_{aj}$ encoding a reaction $aj$ we have the following cases:
\begin{description}
\item[(a)] 
if $aj$ is applicable and it produces the entity $s$,  
the process $P_{aj}$ provides a code of this type:
%By Lemma~\ref{lem:compl1}, all the processes $A_j$, and $Cxt_h$ participated in the tr
%The code executed by any process encoding rule $R_i$ producing $e$ is 
$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\widetilde{s}_i}{\noact}\chainedlink{\noact}{\dots}\link{\widetilde{s}_o}{\dots} \chainend{p_{j+1}}.P_{aj} ;$$

\item[(b)] 
 if $aj$ is applicable and it  consumes the entity $s$,
 the process $P_{aj}$ provides a code of this type: 
$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$

\item[(c)]  
 if $aj$ is applicable and it requires the absence of the entity $s$,
 the process $P_{aj}$  provides a code of this type: 
$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$

\item[(d)] 
 if $aj$ is not applicable,
the process $P_{aj}$ executes a code  capturing either the absence of one of its reactants (case 1), or the presence  of one of its inhibitors (case 2):
\begin{enumerate}
\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{}\link{\overline{s}_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} $;
\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{}\link{s_o}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_{aj}$.
\end{enumerate}
\end{description}
%It is worth nothing that there could be other processes $P_a$ encoding  a reaction $a \in A$ that do not
%involve the entity $s$.\\
Now, we consider the structure of the process $\mathit{Cxt}^1$. 
By Definition~\ref{def:trans}, $Cxt^1$ is the unique process encoding the behaviour of the context regulating the supply of any entity.

\begin{description}
\item[(e)] 
The code of the process $Cxt^1$ that provides $s$ and not $e$ has the following structure: 
$$
\mathit{Cxt}^1 \defeq
\link{\mathit{cxt}}{\dots}
\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{}\link{\widehat{s}_o}{\dots} 
\chainedlink{\underline{e}_i}{\noact}\chainedlink{\noact}{}\link{\underline{e}_o}{\dots} 
\chainend{p_{1}}. \mathit{Cxt}^{2} .
$$
\end{description}

>>>>

 %\begin{proposition}(correctness 1)\\
 %\label{lem:struct}
%  Let ${\cal A} = (S, A)$ be a rs, and let  $\pi=(\gamma,\delta)$ be a $n$-step interactive process in $A$, with
% $\gamma = C_0,\dots,C_n$ and $\delta = D_0,\dots,D_n$.
%% We define $\pi' = (\gamma',\delta')$ with $\gamma' = C_0',\dots,C_{n-1}'$ and $\delta' = D_0',\dots,D_{n-1}'$ as $C_0' = C_1 \cup D_1$,  and
%% $C_i' = C_{i+1}$, $D_{i}' = D_{i+1}$ with $ 1 \leq i < n$.
%  Let   $P=[|{\cal A},\gamma|]$ be  its c\CNA~translation, and let $P'$ be such that $P \xrightarrow{\link{\silent}{\silent}}P'$.% then $P' = [|{\cal A},\gamma' |]$. 
%Then it holds that\\
% $P= \restrict{reacts,ctxs,ents} (\Pi_{a \in A}P_a| \Pi_{s \in S}CXT_s | \Pi_{s \in C_0}P_s| \Pi_{s \notin C_0}\overline{P}_s)$\\
% $P'= \restrict{reacts,ctxs,ents} (\Pi_{a \in A}P_a| \Pi_{s \in S}CXT_s | \Pi_{s \in C_1 \cup D_1}P_s| \Pi_{s \notin C_1 \cup D_1}\overline{P}_s)$
 %\end{proposition}
 \begin{proof}
 First, we note that all the channels in the system are restricted, see Def.~\ref{def:trans}, then
 it holds that the transition labels are of the form  $v=\link{\silent}{\silent}\dots\link{\silent}{\silent}$.\\
 Now,
by Definition~\ref{def:trans} and by Lemma~\ref{lem:struct}.1, all the  channels $r_j,p_j$, with
 $j \in [1,\dots,u]$, and $cxt$,
 %$cxt_h$, with $h \in [1, \dots, w]$,
  and all the annotated versions of 
 $s_i$, $s_o$ are restricted. Also, processes $\mathit{Cxt}$ always requires the 
 interaction with $P_s$ on either on channels $\widehat{s}_i$, $\widehat{s}_o$ or on channels 
 $\underline{s}_i$, $\underline{s}_o$.
 It derives that all the processes: $P_a$  (coding the behaviour of reaction $a \in A$), $P_s$ (coding the behaviour of entity $s \in S$), and $\mathit{Cxt}$ (coding the behaviour of the context regarding all the entities) have been involved in the transition.

For any process $P_{aj}$ encoding a reaction $aj$ we have the following cases:
\begin{description}
\item[(a)] 
if $aj$ is applicable and it produces the entity $s$,  
the process $P_{aj}$ provides a code of this type:
%By Lemma~\ref{lem:compl1}, all the processes $A_j$, and $Cxt_h$ participated in the tr
%The code executed by any process encoding rule $R_i$ producing $e$ is 
$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\chainedlink{\widetilde{s}_i}{\noact}\chainedlink{\noact}{\widetilde{s}_o}\link{}{\dots} \chainend{p_{j+1}}.P_{aj} ;$$

\item[(b)] 
 if $aj$ is applicable and it  consumes the entity $s$,
 the process $P_{aj}$ provides a code of this type: 
$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{s_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$

\item[(c)]  
 if $aj$ is applicable and it requires the absence of the entity $s$,
 the process $P_{aj}$  provides a code of this type: 
$$P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} ;$$

\item[(d)] 
 if $aj$ is not applicable,
the process $P_{aj}$ executes a code  capturing either the absence of one of its reactants (case 1), or the presence  of one of its inhibitors (case 2):
\begin{enumerate}
\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{\overline{s}_i}{\noact}\chainedlink{\noact}{\overline{s}_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}}.P_{aj} $;
\item $P_{aj}\defeq \link{r_j}{\dots}\chainedlink{s_i}{\noact}\chainedlink{\noact}{s_o}\link{}{\dots} \chainedlink{r_{j+1}}{\noact}\chainedlink{\noact}{p_j}\link{}{\dots}\link{}{p_{j+1}} .P_{aj}$.
\end{enumerate}
\end{description}
%It is worth nothing that there could be other processes $P_a$ encoding  a reaction $a \in A$ that do not
%involve the entity $s$.\\
Now, we consider the structure of the process $\mathit{Cxt}^1$. 
By Definition~\ref{def:trans}, $Cxt^1$ is the unique process encoding the behaviour of the context regulating the supply of any entity.

\begin{description}
\item[(e)] 
The code of the process $Cxt^1$ that provides $s$ and not $e$ has the following structure: 
$$
\mathit{Cxt}^1 \defeq
\link{\mathit{cxt}}{\dots}
\chainedlink{\widehat{s}_i}{\noact}\chainedlink{\noact}{\widehat{s}_o}\link{}{\dots} 
\chainedlink{\underline{e}_i}{\noact}\chainedlink{\noact}{}\link{\underline{e}_o}{\dots} 
\chainend{p_{1}}. \mathit{Cxt}^{2} .
$$
\end{description}
